---
title: "Untitled"
output: html_document
---

#Problem 1.
#Using R, generate a random variable X that has 10,000 random uniform numbers from 1 to N, where N can be any number of your choosing greater than or equal to 6.  Then generate a random variable Y that has 10,000 random normal numbers with a mean of ????????????(N+1)/2.  
```{r}
N <- 6
X <- runif(10000, min = 1, max = N)
Mean <- (N+1)/2
Y <- rnorm(10000 , mean = Mean)
```


#Probability.   Calculate as a minimum the below probabilities a through c.  Assume the small letter "x" is estimated as the median of the X variable, and the small letter "y" is estimated as the 1st quartile of the Y variable.  Interpret the meaning of all probabilities.

#5 points           
#a.   P(X>x | X>y)								
```{r}
x <- median(X)
y <- quantile(Y, prob = c(.25))
sum(X > x & X > y)/sum(X > y)
```

#b.  P(X>x, Y>y)
```{r}
sum(X > x &  Y > y)/length(X)
```

#c.  P(X<x | X>y)
```{r}
sum(X < x & X > y)/sum(X>y)
```


#5 points.   Investigate whether P(X>x and Y>y)=P(X>x)P(Y>y) by building a table and evaluating the marginal and joint probabilities.
```{r}
tab <- c(sum(X<x & Y < y),
       sum(X < x & Y == y),
       sum(X < x & Y > y))

tab <- rbind(tab,
              c(sum(X==x & Y < y),
       sum(X == x & Y == y),
       sum(X == x & Y > y))
             
             )
tab <- rbind(tab,
              c(sum(X>x & Y < y),
       sum(X > x & Y == y),
       sum(X > x & Y > y))
             )
tab <- cbind(tab, tab[,1] + tab[,2] + tab[,3])
tab <- rbind(tab, tab[1,] + tab[2,] + tab[3,])
colnames(tab) <- c("Y<y", "Y=y", "Y>y", "Total")
rownames(tab) <- c("X<x", "X=x", "X>x", "Total")
knitr::kable(tab)
```

```{r}
# P(X>x and Y>y)
3723/10000
#P(X>x)P(Y>y)
((5000)/10000)*(7500/10000)
```
#So, P(X > x and Y > y) = 0.37 and P(X>x)P(Y>y) = 0.375. They are almost equal. 

#5 points.  Check to see if independence holds by using Fisher's Exact Test and the Chi Square Test.  What is the difference between the two? Which is most appropriate?
#Fisher's Exact Test
```{r}
fisher.test(table(X>x,Y>y))
```
#The p-value is greater than zero so we cannot reject the null hypothesis because this two events are independent.


#The Chi Square Test
```{r}
chisq.test(table(X>x,Y>y))
```
#The p-value is greeter than zero cannot reject the null hypothesis because this two events are independent.

#Since the p-value for both tests are equivalent and both greater than 0.05 / 5%, we do not reject the null hypothesis of the test.   


#Problem 2
#You are to register for Kaggle.com (free) and compete in the House Prices: Advanced Regression Techniques competition.  https://www.kaggle.com/c/house-prices-advanced-regression-techniques .  I want you to do the following.
```{r}
train <- read.csv('https://raw.githubusercontent.com/Zchen116/Data-605/master/train.csv')
dim(train)
str(train)
```


#5 points.  Descriptive and Inferential Statistics. Provide univariate descriptive statistics and appropriate plots for the training data set.  Provide a scatterplot matrix for at least two of the independent variables and the dependent variable. Derive a correlation matrix for any three quantitative variables in the dataset.  Test the hypotheses that the correlations between each pairwise set of variables is 0 and provide an 80% confidence interval.  Discuss the meaning of your analysis.  Would you be worried about familywise error? Why or why not?
```{r}
numeric <- unlist(lapply(train, is.numeric))  
train_a<-train[ , numeric]
train_b<-subset(train_a, select=-c(Id))
summary(train_b)
```

```{r}
library(DataExplorer)
plot_histogram(train_b)
```

#Now assume we are interested in GrLivArea, and would like to build a model to predict it
```{r}
plot_boxplot(train_b, by = "GrLivArea")
```

```{r}
plot_scatterplot(train_b,  by = "GrLivArea")
```

```{r}
plot_correlation(train_b)
```

#Three selected variables are: SalePrice,TotalBsmtSF,GrLivArea
```{r}
corr_data<-subset(train_b,select=c("SalePrice","TotalBsmtSF", "GrLivArea"))
correlation_matrix <- cor(corr_data)
print(correlation_matrix)
```
#From the Co-relation matrix that we can know 'Saleprice' has strong corelations with 'TotalBsmtSF' and 'GrLivArea' with corelation coefficients of 0.61 and 0.71, and 'TotalBsmtSF' and 'GrLivArea' have moderate corelation between them with coefficient of 0.45.

#Hypothesis testing: the correlations between each pairwise set of variables
#Testing between 'TotalBsmtSF' and 'SalePrice'
```{r}
cor.test(corr_data$TotalBsmtSF, corr_data$SalePrice, method = "pearson", conf.level = 0.8)
```

#Testing between 'GrLivArea' and 'SalePrice'
```{r}
cor.test(corr_data$GrLivArea, corr_data$SalePrice, method = "pearson", conf.level = 0.8)
```

#Testing between 'GrLivArea' and 'TotalBsmtSF'
```{r}
cor.test(corr_data$GrLivArea, corr_data$TotalBsmtSF, method = "pearson", conf.level = 0.8)
```

#In three testings, we have generated an 80 percent confidence interval. We should also note the small p value that we can reject the the null hypothesis and conclude that the true correlation is not 0 for the selected variables. 

#The familywise error rate (FWE or FWER) is the probability of a coming to at least one false conclusion in a series of hypothesis tests . In other words, it's the probability of making at least one Type I Error. The term "familywise" error rate comes from family of tests, which is the technical definition for a series of tests on data.The FWER is also called alpha inflation or cumulative Type I error.           

#Would you be worried about familywise error?
#Yes, of course I would worry about familywise error becuse there are many variables in this dataset that might have impact on the corelation of the the pairs of selected variables that are being tested here. 

#5 points. Linear Algebra and Correlation.  Invert your correlation matrix from above. (This is known as the precision matrix and contains variance inflation factors on the diagonal.) Multiply the correlation matrix by the precision matrix, and then multiply the precision matrix by the correlation matrix. Conduct LU decomposition on the matrix.  
```{r}
print(correlation_matrix)
```

#Invert correlation matrix
```{r}
require(Matrix)
prec_matrix <- solve(correlation_matrix)
print(prec_matrix)
```

#Multiply the correlation matrix by the precision matrix
```{r}
round(correlation_matrix %*% prec_matrix)
```

#Multiply precision matrix by the correlation matrix
```{r}
round(prec_matrix %*% correlation_matrix)
```

#Conduct LU decomposition on the matrix. 
```{r}
lu_mat<-lu(correlation_matrix)
lu_mat2<-expand(lu_mat)
print(lu_mat2$L %*% lu_mat2$U)
```


#5 points.  Calculus-Based Probability & Statistics.  Many times, it makes sense to fit a closed form distribution to data.  Select a variable in the Kaggle.com training dataset that is skewed to the right, shift it so that the minimum value is absolutely above zero if necessary.  Then load the MASS package and run fitdistr to fit an exponential probability density function.  (See  https://stat.ethz.ch/R-manual/R-devel/library/MASS/html/fitdistr.html ).  Find the optimal value of ??? for this distribution, and then take 1000 samples from this exponential distribution using this value (e.g., rexp(1000, ???)).  Plot a histogram and compare it with a histogram of your original variable.   Using the exponential pdf, find the 5th and 95th percentiles using the cumulative distribution function (CDF).   Also generate a 95% confidence interval from the empirical data, assuming normality.  Finally, provide the empirical 5th percentile and 95th percentile of the data.  Discuss.
```{r}
plot_histogram(train$GrLivArea);
summary(train$GrLivArea)
```

```{r}
library(MASS)
fit <- fitdistr(train$GrLivArea,'exponential')
lamda <- fit$estimate
example_distibution <- rexp(1000, lamda)
summary(example_distibution);
hist(example_distibution, main = "Simulated Grade Living Area", xlab="", col = "blue", breaks = 25)
```

#Using the exponential pdf, find the 5th and 95th percentiles using the cumulative distribution function (CDF).
```{r}
quantile(example_distibution, c(.05, .95))
```

```{r}
error <- qnorm(.95) * (fit$sd / sqrt(fit$n))

CI <- data.frame(1 - error, 1 + error)
colnames(CI) <- c("se-","se+")
CI
```

```{r}
quantile(train$GrLivArea, c(.05, .95))
```

#10 points.  Modeling.  Build some type of multiple regression  model and submit your model to the competition board.  Provide your complete model summary and results with analysis.  Report your Kaggle.com user name and score.

## Read test data
```{r}
test <- read.csv('https://raw.githubusercontent.com/Zchen116/Data-605/master/test.csv')
dim(test)
str(test)
```

#we can see categorical variables were converted to numerical values
```{r}
num <- sapply(train, is.numeric)
num_df <- train[ , num]
head(num_df)
```

```{r}
cor_Sales <-data.frame(apply(num_df,2, function(col)cor(col, num_df$SalePrice, use = "complete.obs")))
colnames(cor_Sales) <- c("cor")
cor_Sales
(subset(cor_Sales, cor > 0.5))
```

```{r}
model <- lm(SalePrice ~ OverallQual + YearBuilt + YearRemodAdd + TotalBsmtSF + X1stFlrSF + GrLivArea + FullBath + TotRmsAbvGrd + GarageCars + GarageArea, data = train)
```

```{r}
step_lm <- stepAIC(model, trace=FALSE)
summary(step_lm)
```
#$R^2$ value of 0.7737, 77.37% of the variance can be explained by this model.

```{r}
plot(step_lm$fitted.values, step_lm$residuals, 
     xlab="Fitted Values", ylab="Residuals", main="Fitted Values vs. Residuals")
abline(h=0)
```

#The residuals are normally distributed. 

#Prediction:
```{r}
mySalePrice <- predict(step_lm,test)
prediction <- data.frame( Id = test[,"Id"],  SalePrice = mySalePrice)
prediction[prediction<0] <- 0
prediction <- replace(prediction,is.na(prediction),0)

write.csv(prediction, file="prediction.csv", row.names = FALSE)
```


#Kaggle Submission

#My Kaggle username is **zchen116**. My Score is **0.85356**


